
### Setup
```{r}
library(dplyr)
library(ggplot2)
library(readxl)
library(tidyverse)

def <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Defensive')
off <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Offensive')
kikr <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Kick Returns')
kik <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Kicking')
bio <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Biography')
```


<h1>Part 1 Question 1</h1>
```{R}

def$Name <- factor(def$Name)
def$Opponent_Opponent <- factor(def$Opponent_Opponent)
off$Name <- factor(off$Name)
off$Opponent_Opponent <- factor(off$Opponent_Opponent)


```


<h1>Part 1 Question 2</h1>
```{r}

defClean <- def %>%
  mutate(Tackles_Solo=as.numeric(Tackles_Solo)) %>%
  mutate(Tackles_ASST=as.numeric(Tackles_ASST)) %>%
  mutate(Tackles_TFL=as.numeric(Tackles_TFL)) %>%
  mutate(Tackles_Sack=as.numeric(Tackles_Sack)) %>%
  mutate(Turnover_FF=as.numeric(Turnover_FF)) %>%
  mutate(Turnover_FR=as.numeric(Turnover_FR)) %>%
  mutate(Turnover_INT=as.numeric(Turnover_INT)) %>%
  mutate(Pass_QBH=as.numeric(Pass_QBH)) %>%
  mutate(Pass_PB=as.numeric(Pass_PB))
  
offClean <- off %>%
  mutate(Receiving_REC=as.numeric(Receiving_REC)) %>%
  mutate(Receiving_YDS=as.numeric(Receiving_YDS)) %>%
  mutate(Receiving_TD=as.numeric(Receiving_TD)) %>%
  mutate(Rushing_ATT=as.numeric(Rushing_ATT)) %>%
  mutate(Rushing_YDS=as.numeric(Rushing_YDS)) %>%
  mutate(Rushing_TD=as.numeric(Rushing_TD)) %>%
  mutate(`Passing_CMP-ATT`=as.numeric(`Passing_CMP-ATT`)) %>%
  mutate(Passing_YDS=as.numeric(Passing_YDS)) %>%
  mutate(Passing_TD=as.numeric(Passing_TD)) %>%
  mutate(Passing_INT=as.numeric(Passing_INT))
  
bioClean <- bio %>%
  mutate(Weight=as.numeric(Weight))

#str(defClean)
#str(bioClean)
#str(offClean)
```

<h1>Part 1 Question 3</h1>
```{r}
bioClean <- bioClean %>%
  separate(Height, c('Height_feet', 'Height_inches')) %>%
  mutate(Height_feet = as.integer(Height_feet), 
         Height_inches = as.integer(Height_inches)) %>%
  mutate(Height = (Height_inches + (Height_feet * 12))) %>%
  select(-Height_inches, -Height_feet)
```

```{R}
str(defClean)
str(offClean)
str(bioClean)
```

<h1>Part 2 Question 1</h1>
```{R}
offClean <- offClean %>% 
  pivot_longer(cols = Receiving_REC:Passing_INT, names_to = "stat")
```

<h1>Part 2 Question 2</h1>
```{R}
ggplot(data = offClean, aes(value)) +
  geom_histogram(color = "steelblue") + 
  labs(title = "Distribution of Offensive Statistics",
       y = "Number of Occurrences", x = "Value of Statistic") + 
  facet_wrap(~ stat)
```

<h1>Part 2 Question 3</h1>
```{r}
okoff <- offClean %>%
  filter(Opponent_Opponent == 'Oklahoma') %>%
  filter(stat == 'Receiving_YDS')
okoff
oroff <- offClean %>%
  filter(Opponent_Opponent == "Oregon") %>%
  filter(stat == 'Receiving_YDS')
okoroff <- oroff %>%
  full_join(okoff)
ggplot(oroff, aes(value)) + geom_dotplot() + xlab('Number of Receiving Yards Against Oregon') + ylab('Proportion of Players') + ggtitle('Receiving Yards Against Oregon Where Each Dot Represents a Player')
ggplot(okoff, aes(value)) + geom_dotplot() + xlab('Number of Receiving Yards Against Oklahoma') + ylab('Proportion of Players') + ggtitle('Receiving Yards Against Oklahoma Where Each Dot Represents a Player')
```

<h1>Part 2 Question 4</h1>
```{r}
bioClean <- bioClean %>%
  separate(Hometown, c('Hometown_city', 'Hometown_state'), sep = ', ')
bioClean %>%
  select(Hometown_city, Hometown_state) %>%
  head()
```

<h1>Part 2 Question 5</h1>
```{r}
bioClean %>%
  select(Hometown_state) %>%
  group_by(Hometown_state) %>%
  summarise(count = n())
```

<h1>Part 3 Question 1</h1>
```{r}
ggplot(bioClean, aes(Hometown_state)) + geom_bar()
```

<h1>Part 3 Question 2</h1>
Not surprisingly, the amount of touchdowns the team makes is dependent to the amount of touchdowns that the quarterback (Brock Purdy) makes.
```{r}
defense20 <- read_excel('cyclonesFootball2020.xlsx', sheet='Defensive')
offense20 <- read_excel('cyclonesFootball2020.xlsx', sheet='Offensive')
biography20 <- read_excel('cyclonesFootball2020.xlsx', sheet='Biography')

defense19 <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
offense19 <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
biography19 <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')


offClean20 <- offense20 %>%
  mutate(Receiving_REC=as.numeric(Receiving_REC)) %>%
  mutate(Receiving_YDS=as.numeric(Receiving_YDS)) %>%
  mutate(Receiving_TD=as.numeric(Receiving_TD)) %>%
  mutate(Rushing_ATT=as.numeric(Rushing_ATT)) %>%
  mutate(Rushing_YDS=as.numeric(Rushing_YDS)) %>%
  mutate(Rushing_TD=as.numeric(Rushing_TD)) %>%
  mutate(Passing_YDS=as.numeric(Passing_YDS)) %>%
  mutate(Passing_TD=as.numeric(Passing_TD)) %>%
  mutate(Passing_INT=as.numeric(Passing_INT))

defClean19 <- defense19 %>%
  mutate(Tackles_Solo=as.numeric(Tackles_Solo)) %>%
  mutate(Tackles_ASST=as.numeric(Tackles_ASST)) %>%
  mutate(Tackles_TFL=as.numeric(Tackles_TFL)) %>%
  mutate(Tackles_Sack=as.numeric(Tackles_Sack)) %>%
  mutate(Turnover_FF=as.numeric(Turnover_FF)) %>%
  mutate(Turnover_FR=as.numeric(Turnover_FR)) %>%
  mutate(Turnover_INT=as.numeric(Turnover_INT)) %>%
  mutate(Pass_QBH=as.numeric(Pass_QBH)) %>%
  mutate(Pass_PB=as.numeric(Pass_PB))
  
offClean19 <- offense19 %>%
  mutate(Receiving_REC=as.numeric(Receiving_REC)) %>%
  mutate(Receiving_YDS=as.numeric(Receiving_YDS)) %>%
  mutate(Receiving_TD=as.numeric(Receiving_TD)) %>%
  mutate(Rushing_ATT=as.numeric(Rushing_ATT)) %>%
  mutate(Rushing_YDS=as.numeric(Rushing_YDS)) %>%
  mutate(Rushing_TD=as.numeric(Rushing_TD)) %>%
  mutate(Passing_YDS=as.numeric(Passing_YDS)) %>%
  mutate(Passing_TD=as.numeric(Passing_TD)) %>%
  mutate(Passing_INT=as.numeric(Passing_INT))

defClean20 <- defense20 %>%
  mutate(Tackles_Solo=as.numeric(Tackles_Solo)) %>%
  mutate(Tackles_ASST=as.numeric(Tackles_ASST)) %>%
  mutate(Tackles_TFL=as.numeric(Tackles_TFL)) %>%
  mutate(Tackles_Sack=as.numeric(Tackles_Sack)) %>%
  mutate(Turnover_FF=as.numeric(Turnover_FF)) %>%
  mutate(Turnover_FR=as.numeric(Turnover_FR)) %>%
  mutate(Turnover_INT=as.numeric(Turnover_INT)) %>%
  mutate(Pass_QBH=as.numeric(Pass_QBH)) %>%
  mutate(Pass_PB=as.numeric(Pass_PB))
  
offClean20 <- offense20 %>%
  mutate(Receiving_REC=as.numeric(Receiving_REC)) %>%
  mutate(Receiving_YDS=as.numeric(Receiving_YDS)) %>%
  mutate(Receiving_TD=as.numeric(Receiving_TD)) %>%
  mutate(Rushing_ATT=as.numeric(Rushing_ATT)) %>%
  mutate(Rushing_YDS=as.numeric(Rushing_YDS)) %>%
  mutate(Rushing_TD=as.numeric(Rushing_TD)) %>%
  mutate(Passing_YDS=as.numeric(Passing_YDS)) %>%
  mutate(Passing_TD=as.numeric(Passing_TD)) %>%
  mutate(Passing_INT=as.numeric(Passing_INT))


purdy <- offClean20 %>%
  filter(Name=="Purdy, Brock")
head(purdy)

purdy_summary <- purdy %>% 
  summarise(
    Total_Pass_TD=sum(Passing_TD),
    Total_Rush_TD=sum(Rushing_TD),
    Mean_Pass_TD=mean(Passing_TD),
    Mean_Rush_TD=mean(Rushing_TD))
head(purdy_summary)

team <- offClean20 %>%
  filter(Name!="Purdy, Brock")
head(team)

team_summary <- team %>%
  group_by(Name) %>%
  summarize(
    Total_Receiving_TD=sum(Receiving_TD),
    Total_Rushing_TD=sum(Rushing_TD),
    Mean_Receiving_TD=mean(Receiving_TD),
    Mean_Rushing_TD=mean(Rushing_TD)
  )
head(team_summary)
```

<h1>Part 3 Question 3</h1>
JaQuan Bailey went from 10 total TD to 58. This is a large improvement from last year arguably making him the most improved player.
```{r}

def_summary19 <- defClean19 %>%
  group_by(Name) %>%
  mutate(Tackles_Total=Tackles_Solo+Tackles_ASST+Tackles_TFL+Tackles_Sack) %>%
  summarize(
    Total_Tackles=sum(Tackles_Total),
    Mean_Total_Tackles=mean(Tackles_Total),
    Mean_Tackles_Solo=mean(Tackles_Solo),
    Mean_Tackles_ASST=mean(Tackles_ASST),
    Mean_Tackles_TFL=mean(Tackles_TFL),
    Mean_Tackles_Sack=mean(Tackles_Sack),
    Mean_Turnover_FF=mean(Turnover_FF),
    Mean_Turnover_FR=mean(Turnover_FR),
    Mean_Turnover_INT=mean(Turnover_INT),
  )
head(def_summary19)

def_summary20 <- defClean20 %>%
  group_by(Name) %>%
  mutate(Tackles_Total=Tackles_Solo+Tackles_ASST+Tackles_TFL+Tackles_Sack) %>%
  summarize(
    Total_Tackles=sum(Tackles_Total),
    Mean_Total_Tackles=mean(Tackles_Total),
    Mean_Tackles_Solo=mean(Tackles_Solo),
    Mean_Tackles_ASST=mean(Tackles_ASST),
    Mean_Tackles_TFL=mean(Tackles_TFL),
    Mean_Tackles_Sack=mean(Tackles_Sack),
    Mean_Turnover_FF=mean(Turnover_FF),
    Mean_Turnover_FR=mean(Turnover_FR),
    Mean_Turnover_INT=mean(Turnover_INT),
  )
head(def_summary20)
#Shows JaQuan Bailey went from 10 total TD to 58

str(offClean20)

off_summary19 <- offClean19 %>%
  group_by(Name) %>%
  summarize(
    Total_Receiving_TD=sum(Receiving_TD),
    Total_Rushing_TD=sum(Rushing_TD),
    Mean_Receiving_TD=mean(Receiving_TD),
    Mean_Rushing_TD=mean(Rushing_TD)
  )
head(off_summary19)

off_summary20 <- offClean20 %>%
  group_by(Name) %>%
  summarize(
    Total_Receiving_TD=sum(Receiving_TD),
    Total_Rushing_TD=sum(Rushing_TD),
    Mean_Receiving_TD=mean(Receiving_TD),
    Mean_Rushing_TD=mean(Rushing_TD)
  )
head(off_summary20)
```

